Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-11-08 paper/article.tex
Paper progress

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{graphicx}
\graphicspath{ {./figs/} }
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\begin{document}

@@

The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
consists of a ``modular'' vanilla RNN, and a linear readout layer. The inputs are visual
features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
the same objects which the monkeys grasped. The outputs are muscle length velocities for the
should, arm, and hand of the monkey during the trial. The natural velocities were captured
with a motion capturing glove, and the artificial neural network is trained to recapitulate
those grasping motions. See Fig. \ref{fig:michaels}.

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once train, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. Notably, the activity of
the module receiving the visual inputs resembles the AIP activity of the monkey subject
from the same trials. Likewise, the second and third modules resemble the natural activity
from the monkey's F5 and M1 regions, respectively. The network dynamics show a number of
other correspondences to the monkeys' natural brain activity as well, as detailed in the paper.

Notably, the simulated circuit's activity shows a relatively clear separation of the
object shapes.  That is - the visual information input to the network differentiates one trial
from another, and the circuit is able to leverage that information to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and location.
The visual information forward-propagates through the network, through successive processing
steps, where it is eventually leveraged to recapitulate the grasp appropriate for the object
represented by the input visual features. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

As a result, simulating a brain lesion in terms of this artificial network results in error modes
which resemble a natural lesion of certain regions of a primate brain.

...
* Lesioning properties

@@

* Show example where AIP/F5 lesioned and how that affects performance

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%    INSTITUTE OF PHYSICS PUBLISHING                                   %
%                                                                      %
%   `Preparing an article for publication in an Institute of Physics   %
%    Publishing journal using LaTeX'                                   %
%                                                                      %
%    LaTeX source code `ioplau2e.tex' used to generate `author         %
%    guidelines', the documentation explaining and demonstrating use   %
%    of the Institute of Physics Publishing LaTeX preprint files       %
%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
%                                                                      %
%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
% First we have a character check
%
% ! exclamation mark    " double quote  
% # hash                ` opening quote (grave)
% & ampersand           ' closing quote (acute)
% $ dollar              % percent       
% ( open parenthesis    ) close paren.  
% - hyphen              = equals sign
% | vertical bar        ~ tilde         
% @ at sign             _ underscore
% { open curly brace    } close curly   
% [ open square         ] close square bracket
% + plus sign           ; semi-colon    
% * asterisk            : colon
% < open angle bracket  > close angle   
% , comma               . full stop
% ? question mark       / forward slash 
% \ backslash           ^ circumflex
%
% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
% abcdefghijklmnopqrstuvwxyz 
% 1234567890
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[12pt]{iopart}
\usepackage{graphicx,epstopdf}
\usepackage{caption}
\usepackage{subcaption}
\graphicspath{ {./figs/} }
%Uncomment next line if AMS fonts required
%\usepackage{iopams}
\begin{document}

@@

The simulated circuit, from Michaels et al. \cite{michaels.mrnn, susillo.mrnn}, was trained
to resemble the grasping circuits of monkey subjects engaged in a reach-to-grasp task. It
consists of a ``modular'' vanilla RNN (mRNN), and a linear readout layer. The inputs are visual
features intended to capture the view the monkeys had during the task, i.e. 3D renderings of
the same objects which the monkeys grasped. The outputs are muscle length velocities for the
should, arm, and hand of the monkey during the trial. The natural velocities were captured
with a motion capturing glove, and the artificial neural network is trained to recapitulate
those grasping motions. Data and trained models from this work were supplied to us by the lead
author Jonathan Michaels, for the purpose of our present simulation. See Fig. \ref{fig:michaels}.

\begin{figure}
\includegraphics[width=\textwidth]{michaels.eps}
\caption{Michaels Modular RNN (mRNN), a simulated grasping circuit. The emergent dynamics of the
three modules correspond well to natural neural activity measured from primate AIP, F5, M1 regions,
respectively, from the same task. Visual (VGGNet) features forward propagate through the
network, conditioning the grasp for the object of the particular size, shape, and location.}
\centering
\label{fig:michaels}
\end{figure}

The design of the circuit is intended to resemble a vision-to-grasp pipeline, essentially
representing the visual processing needed to reach the hand to the appropriate position for
the grasp, and to form the hand properly for grasping the particular shape of object. The
emergent dynamics of the artificial network's ``modules'', once trained, correspond roughly
to the AIP, F5, and M1 portions of the monkey subjects' brains. `Correspond' here, notably,
refers to the fact that the activity of the module receiving the visual inputs resembles
the AIP activity of the monkey subject from the same trials. Likewise, the second and
third modules resemble the natural activity from the monkey's F5 and M1 regions, respectively.
The emergent network dynamics show a number of other correspondences to the monkeys' natural
brain activity as well, as detailed in the paper.

Importantly, the simulated circuit's activity shows a relatively clear separation of the
object shapes. That is - the visual information input to the network differentiates one trial
from another, and the circuit is able to leverage that information to properly condition
the hand shape trajectory for grasping the object of that trial's particular shape, size, and
location. The visual information forward-propagates through the network, through successive
processing steps, where it is eventually leveraged to recapitulate the grasp appropriate
for the object represented by the input visual features. It follows, then, that the classes
are separable in some way by observing the distinct ways that the mRNN is activated by the
classes' corresponding visual inputs. As we will note below: in the absense of this
visual information forward propagating, the circuit can at-best learn a loss-minimizing grasp,
stereotyped across all object sizes and shapes. Such a grasp bears some resemblance to all
grasps the given monkey performed, due to all trials being a reach-to-grasp preceded by a
waiting period, but performance suffers drastically.

As a result, simulating a brain lesion in terms of this artificial network results in error
modes which resemble a natural lesion of certain regions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate visual object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. Likewise, if we simulate a
``stroke'' of a portion of the output module of the Michaels mRNN, roughly
corresponding to M1, we see a more wholesale loss of movement, affecting even the
ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.46\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
	    \caption{inner1}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.46\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
	    \caption{inner1}
	\end{subfigure}
	\hfill
	\caption{Two simple graphs}
	\label{fig:lesion}
\end{figure}

@@
